Downscale t2m for KSS

### 2016-10-17 EU-Circle/eu-circle_case-study-3.Rmd
moved ABC4CDE to own repository

```{r,echo=TRUE}  
## Load the necessary R-package
library(esd)
## Information about the system and session
sessionInfo()
```

---------------------------------
```{r,eval=FALSE}
## Get MIDAS preciptation data that has laready been retrieved 
## from the MIDAS data base (UK Met Office)
load('pr.eu-circle-torbay.rda')
```

@@

```{r}
PrexpPr <- function(mu,fw,x0=10) {
  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))

  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
  attr(Pr,'prob.f(mu)') <- Pr.mu
  attr(Pr,'prob.f(fw)') <- Pr.fw
  Pr <- attrcp(mu,Pr)
  attr(Pr,'variable') <- 'Pr'
  attr(Pr,'unit') <- 'probability'
  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
  class(Pr) <- class(mu)
  invisible(Pr)
}
```

@@

```{r}
load('pr.eu-circle-torbay.rda') 
pr <- subset(pr,it=c(1960,2013))

@@

@@

```{r}
 ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
pr.ecad <- station(ss)
mu.ecad <- subset(annual(pr.ecad,'wetmean',nmin=100),it=c(1961,2011))
fw.ecad <- subset(annual(pr.ecad,'wetfreq',nmin=100),it=c(1961,2011))
## Second weeding of stations with many missing values in annual 
## aggregate
nok <- apply(coredata(mu.ecad),2,'nv')
mu.ecad <- subset(mu.ecad,is=(nok>= 50))
fw.ecad<- subset(fw.ecad,is=(nok>= 50))
diagnose(mu.ecad)
```

@@

```{r}
mu <- combine.stations(mu,mu.ecad)
fw <- combine.stations(fw,fw.ecad)
```

```{r}
##Fill in the remaining gaps using a PCA-based interpolation scheme 
## Check the records
diagnose(mu)
pcafill(mu) -> mu
pcafill(fw) -> fw
```

@@

```{r}
Pr.gt.10 <- PrexpPr(mu,fw,x0=10)
plot(Pr.gt.10); grid()
for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.10[,i]))
```

@@

```{r}
## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
MU <- aggregate(x,month,FUN='wetmean')
## Estimate the mean seasonal cycle in wet-day mean frequency
FW <- aggregate(x,month,FUN='wetfreq')
## Estimate the mean seasonal cycle in wet-day mean frequency
MP <- aggregate(x,month,FUN='mean')
```

```{r}
## Show the mean seasonal cycle in the wet-day mean precipitation
plot(MU)
```

@@

```{r}
## Show the mean seasonal cycle in the wet-day frequency
plot(FW)
```

@@

```{r}
## Show the mean seasonal cycle in the total precipitation
MP <- 30*MP
attr(MP,'unit') <- 'mm/month'
plot(MP)
```

@@

```{r}
## Estimate the PCA for the annual cycle to identify the general features:
pca.MU <- PCA(MU,n=10)
plot(pca.MU)
```

@@

```{r}
if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
  dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE,verbose=TRUE)
  save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
} else load('dse.mu.eu-circle-worstcase.rda')
plot(dse.mu[[1]])
```

@@

```{r}
index(newlyn) <- year(newlyn)
index(sonel) <- year(sonal)
index(gloss) <- year(gloss)
Z <- combine.stations(newlyn,sonel,gloss)
nv <- apply(coredata(Z),2,FUN='nv')
Z <- subset(Z,is=(nv > 50))
Z <- subset(Z,it=c(1950,2015))
plot(Z); grid()
```

@@

```{r slp-trend-map}
slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v12.0.nc',lon=c(-6,0),lat=c(48,52))
slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
attr(slp.eobs,'unit') <- 'Pa'
map(slp.eobs,FUN='trend')
```

@@

```{r predictand}
pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
mu <- annual(pr.eobs,FUN='wetmean')
fw <- annual(pr.eobs,FUN='wetfreq')
```

>>>>

```{r,echo=TRUE}  
## Load the necessary R-package
library(esd)
## Information about the system and session
```

---------------------------------

@@

```{r function-probability}
PrexpPr <- function(mu,fw,x0=10) {
  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))

  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
  attr(Pr,'prob.f(mu)') <- Pr.mu
  attr(Pr,'prob.f(fw)') <- Pr.fw
  Pr <- attrcp(mu,Pr)
  attr(Pr,'variable') <- 'Pr'
  attr(Pr,'unit') <- 'probability'
  attr(Pr,'longname') <- paste('Probability of exceeding',x0,'mm')
  class(Pr) <- class(mu)
  invisible(Pr)
}
```

@@

```{r read-precip-data}
## Get MIDAS preciptation data that has laready been retrieved 
## from the MIDAS data base (UK Met Office)
load('pr.eu-circle-torbay.rda') 
pr <- subset(pr,it=c(1960,2013))

@@

## Quality check: remove outliers days with more than 300 mm - set to NA
xc <- coredata(x); xc[xc > 300] <- NA; xc -> coredata(x); rm('xc')

@@

```{r ecad-data}
 ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
pr.ecad <- station(ss)
mu.ecad <- subset(annual(pr.ecad,'wetmean',nmin=100),it=c(1961,2011))
fw.ecad <- subset(annual(pr.ecad,'wetfreq',nmin=100),it=c(1961,2011))
## Second weeding of stations with many missing values in annual 
## aggregate
nok <- apply(coredata(mu.ecad),2,'nv')
mu.ecad <- subset(mu.ecad,is=(nok>= 50))
fw.ecad<- subset(fw.ecad,is=(nok>= 50))
diagnose(mu.ecad)
```

@@

```{r combine-stations}
mu <- combine.stations(mu,mu.ecad)
fw <- combine.stations(fw,fw.ecad)
```

```{r fill in missing data}
##Fill in the remaining gaps using a PCA-based interpolation scheme 
## Check the records
diagnose(mu)
pcafill(mu) -> mu
pcafill(fw) -> fw
```

@@

```{r probabilities}
Pr.gt.30 <- PrexpPr(mu,fw,x0=30)
plot(Pr.gt.30); grid()
for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.30[,i]))
```

@@

```{r aggregate}
## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
MU <- aggregate(x,month,FUN='wetmean')
## Estimate the mean seasonal cycle in wet-day mean frequency
FW <- aggregate(x,month,FUN='wetfreq')
## Estimate the mean seasonal cycle in wet-day mean frequency
MP <- aggregate(x,month,FUN='mean')
```

```{r precipitation-intensity}
## Show the mean seasonal cycle in the wet-day mean precipitation
plot(MU)
```

@@

```{r wet-day frequency}
## Show the mean seasonal cycle in the wet-day frequency
plot(FW)
```

@@

```{r sesonal-precipitation}
## Show the mean seasonal cycle in the total precipitation
MP <- 30*MP
attr(MP,'unit') <- 'mm/month'
plot(MP)
```

@@

```{r winter-rain}
## Estimate the PCA for the annual cycle to identify the general features:
pca.MU <- PCA(MU,n=10)
plot(pca.MU)
```

@@

```{r downscale-ensemble}
if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
  dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
                                lon=c(-80,20),lat=c(10,60),
                                rel.cord=FALSE,verbose=TRUE)
  save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
} else load('dse.mu.eu-circle-worstcase.rda')
#plot(dse.mu[[1]])
plot.zoo(dse.mu[[1]],plot.type='single',col=rgb(1,0,0,0.05))
grid()
```

@@

```{r}
index(newlyn) <- year(newlyn)
index(sonel) <- year(sonel)
index(gloss) <- year(gloss)
Z <- combine.stations(newlyn,sonel,gloss)
nv <- apply(coredata(Z),2,FUN='nv')
Z <- subset(Z,is=(nv > 50))
Z <- subset(Z,it=c(1950,2015))
plot(Z); grid()
```

@@

```{r slp-trend-map}
slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v13.1.nc',lon=c(-6,0),lat=c(48,52))
slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
attr(slp.eobs,'unit') <- 'Pa'
map(slp.eobs,FUN='trend')
```

@@

```{r predictand}
pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v13.0.nc',lon=c(-6,-1),lat=c(50,52))
mu <- annual(pr.eobs,FUN='wetmean')
fw <- annual(pr.eobs,FUN='wetfreq')
```

### 2016-10-17 KSS/ds-pr4KSS.Rmd
moved ABC4CDE to own repository

```{r}
## Aggregate sub-season values to annual mean values
if (!exists("es.c")) {
  slp <- retrieve('slp.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  slp.w <- annual(subset(slp,it=month.abb[4:9]),'mean',nmin=6)
  slp.c <- annual(subset(slp,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  t2m <- retrieve('air.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  t2m.w <- annual(subset(t2m,it=month.abb[4:9]),'mean',nmin=6)
  t2m.c <- annual(subset(t2m,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.w <- annual(subset(C.C.eq(t2m),it=month.abb[4:9]),'mean',nmin=6)
  es.w <- mask(es.w,land=TRUE)
  es.c <- annual(subset(C.C.eq(t2m),it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.c <- mask(es.w,land=TRUE)
}
## Represent the predictors in terms of EOFs
eof.t2m.w <- EOF(t2m.w)
eof.t2m.c <- EOF(t2m.c)
eof.slp.w <- EOF(slp.w)
eof.slp.c <- EOF(slp.c)
eof.es.w <- EOF(es.w)
eof.es.c <- EOF(es.c)
```

>>>>

```{r}
## Aggregate sub-season values to annual mean values
if (!exists("es.c")) {
  slp <- retrieve('slp.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  slp.w <- annual(subset(slp,it=month.abb[4:9]),'mean',nmin=6)
  slp.c <- annual(subset(slp,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  t2m <- retrieve('air.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  t2m.w <- annual(subset(t2m,it=month.abb[4:9]),'mean',nmin=6)
  t2m.c <- annual(subset(t2m,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.w <- annual(subset(C.C.eq(t2m),it=month.abb[4:9]),'mean',nmin=6)
  es.w <- mask(es.w,land=TRUE)
  es.c <- annual(subset(C.C.eq(t2m),it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.c <- mask(es.w,land=TRUE)
}
## Represent the predictors in terms of EOFs
eof.t2m.w <- EOF(t2m.w)d
eof.t2m.c <- EOF(t2m.c)
eof.slp.w <- EOF(slp.w)
eof.slp.c <- EOF(slp.c)
eof.es.w <- EOF(es.w)
eof.es.c <- EOF(es.c)
```
